---
title: 'Homework 4: Bags, Forests, Boosts, oh my'
author: "Alex Boyce"
date: "3/6/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message = F, warning = F, include = F}
library(tidyverse)
library(broom)
library(glmnet)
library(caret)
library(ISLR)
library(janitor)
library(stringr)
library(rpart)
library(rpart.plot)
library(partykit)
library(randomForest)
library(MASS)
library(gbm)
theme_set(theme_bw())
```

## Problem 1

Problem 7 from Chapter 8 in the text. To be specific, please use a sequence of
`ntree` from 25 to 500 in steps of 25 and `mtry` from 3 to 9 for by 1. 

## Answer 1
```{r}
set.seed(1)
df <- tbl_df(Boston)
inTraining <- createDataPartition(df$medv, p = .75, list = F)
training <- df[inTraining, ]
testing  <- df[-inTraining, ]

set.seed(1)
results <- tibble(trees = rep(seq(25,500,by=25), each=7),
                  mtry = rep(3:9, each=1,times=20),
                  MSE = rep(NA, length(trees)))


set.seed(1)
for (i in seq(25, 500, by=25)){
  rf_results <- train(medv ~ ., 
                      data = training,
                      method = "rf",
                      ntree = i,
                      importance = T,
                      tuneGrid = data.frame(mtry = seq(3,9,by=1)))
  for (j in seq(1,7)){
    results[which(results$trees==i & results$mtry==j+2),'MSE'] = (rf_results$results$RMSE[j])^2
    }
}

p <- ggplot(data = results, aes(x = trees, y = MSE, group=mtry, col=as.factor(mtry)))
p + geom_line() + labs(color = 'mtry')
```

RMSE fluctuates inconsistently based on the number of trees - 50 looks like a good compromise between time to run and low error. In terms of mtry, RMSE improvement appears to stop at around 5 so that seems like a good option.

####Answer 1 w/o cross-validation
```{r}
#without cross validation
set.seed(1)
results <- tibble(trees = rep(seq(25,500,by=25), each=7),
                  mtry = rep(3:9, each=1,times=20),
                  MSE = rep(NA, length(trees)))

set.seed(1)
for (i in seq(25, 500,by=25)){
  for (k in seq(3, 9, by=1)){
    rf_results <- randomForest(medv ~ ., data = training, ntree = i, mtry=k)
    test_pred <- predict(rf_results, newdata = testing)
    error <- (mean((test_pred - testing$medv)^2))
    results[which(results$trees==i & results$mtry==k),'MSE'] = error
  }
}

p <- ggplot(data = results, aes(x = trees, y = MSE, group=mtry, col=as.factor(mtry)))
p + geom_line() + labs(color = 'mtry')
```


## Problem 2
Problem 8 from Chapter 8 in the text. Set your seed with 9823 and split into 
train/test using 50\% of your data in each split. In addition to 
parts (a) - (e), do the following:

## Answer 2

#### Part a - splitting the data
```{r}
df <- tbl_df(Carseats)

set.seed(9823)
inTraining <- createDataPartition(df$Sales, p = .5, list = F)
training <- df[inTraining, ]
testing  <- df[-inTraining, ]
```


#### Part b - regression tree
```{r}
tree_carseats <- rpart(Sales ~ . , training)
prp(tree_carseats)

test_pred <- predict(tree_carseats, newdata = testing)
mean((test_pred - testing$Sales)^2)
```


#### Part c - Tree cross-validation for tree depth
```{r}
set.seed(9823)
fit_control <- trainControl(method = "repeatedcv", number = 10, repeats = 10)
cv_carseats_tree <- train(Sales ~ ., training, 
                          method = "rpart2", 
                          trControl = fit_control, 
                          tuneGrid = data.frame(maxdepth = 1:10))
plot(cv_carseats_tree)
```

Maxdepth of 3 gives the lowest RMSE - will re-fit using that
```{r}
set.seed(9823)
cv_carseats_tree3 <- rpart(Sales ~ ., training, maxdepth = 3)
prp(cv_carseats_tree3)

test_pred <- predict(cv_carseats_tree3, newdata = testing)
mean((test_pred - testing$Sales)^2)
```


#### Part d - Bagging
```{r}
set.seed(9823)
carseats_bag <- randomForest(Sales ~ ., data = training, 
                             mtry = ncol(training)-1,
                             ntrees=500,
                             importance=TRUE)

test_pred <- predict(carseats_bag, newdata = testing)
mean((test_pred - testing$Sales)^2)
```

Plotting variable importance
```{r}
imp <- varImp(carseats_bag)
rn <- row.names(imp)
imp_df <- tibble(variable = rn, importance = imp$Overall) %>%
  arrange(desc(-importance)) %>%
  mutate(variable = factor(variable, variable))

p <- ggplot(data = imp_df, aes(variable, importance))
p + geom_col(fill = "#6e0000") +
  coord_flip()
```
Shelf location and price are the two most important variables here


#### Part e - Random Forest
```{r}
set.seed(9823)
carseats_rf <- train(Sales ~ ., data = training,
                     method = "rf", ntree = 250,
                      importance = T, tuneGrid = data.frame(mtry = 2:10))

p <- ggplot(data = carseats_rf$results, aes(x = mtry, y = RMSE))
p + geom_point() +
  geom_line()
```

mtry of 5-8 results in similiar RMSE - will use 5 for simplicity

```{r}
set.seed(9823)
carseats_rf5 <- randomForest(Sales ~ ., data = training, mtry = 5, importance = TRUE)

test_pred <- predict(carseats_rf5, newdata = testing)
mean((test_pred - testing$Sales)^2)
```

Variable Importance
```{r}
imp <- varImp(carseats_rf5)
rn <- row.names(imp)
imp_df <- tibble(variable = rn, importance = imp$Overall) %>%
  arrange(desc(-importance)) %>%
  mutate(variable = factor(variable, variable))

p <- ggplot(data = imp_df, aes(variable, importance))
p + geom_col(fill = "#6e0000") +
  coord_flip()
```
Again, shelf location and price are most important


#### 1. Fit a gradient-boosted tree to the training data and report the estimated test MSE. 
```{r}
set.seed(9823)
grid <- expand.grid(interaction.depth = c(1, 3), 
                    n.trees = seq(0, 2000, by = 100),
                    shrinkage = c(.01, 0.001),
                    n.minobsinnode = 10)

trainControl <- trainControl(method = "cv", number = 5)
carseats_gb <- train(Sales ~ ., 
                     data = training,
                     distribution = "gaussian", 
                     method = "gbm",
                     trControl = trainControl, 
                     tuneGrid = grid,
                     verbose = FALSE)

test_pred <- predict(carseats_gb, newdata = testing)
mean((test_pred - testing$Sales)^2)
```


#### 2. Fit a multiple regression model to the training data and report the estimated test MSE
```{r}
carseats_lm <- lm(Sales ~ .,data = training)
test_pred <- predict(carseats_lm, newdata = testing)
mean((test_pred - testing$Sales)^2)
```


#### 3. Summarize your results. 
Tree error MSE steadily improved throughout the problem, with the original tree & cross-validation of depth around 4.5-5. Bagging and random forest dropped MSE to around 3, then boosting further dropped to 1.8. However, a simple multiple regression with all variables was still best at 1.01 and is probably the easiest to explain.
